---
title: "p8105_hw3_ar4459_practice"
author: "ASHLEY ROMO"
date: "2023-10-05"
output: github_document
---

## Problem 1

```{r}
library(tidyverse)
library(p8105.datasets)
data("instacart")
```

There are a total of 1,384,617 rows and 15 columns. The structure of the data is contains many ids including order_id, product_id, user_id, aisle_id, and department_id. Some key variables include product names, order_hour_of_day, and reordered because they describe at what time during the day specific products are purchased or reordered, which can help the store advertise the most sold items during the hours they are most often purchased or reordered.

```{r}
instacart_df = instacart

aisles_df = instacart_df |> 
  select(order_id, aisle_id, aisle) 
  


```







# Problem 2

First, I load the BRFSS data.

```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

Now, I will do some data cleaning

```{r}
brfss_df = brfss_smart2010 |> 
  janitor::clean_names() |> 
  select( year, loc_abbr = locationabbr, loc_des = locationdesc, everything()) |> 
  filter(response == "Excellent" | response == "Poor")
```






## Problem 3

First, load the data

```{r}
accel_df = read_csv("data/nhanes_accel.csv") |> 
  janitor::clean_names() |> 
  pivot_longer(
    min1:min1440, 
    names_to = "min",
    values_to = "phys_act"
  )


covar_df = read_csv("data/nhanes_covar.csv", skip = 4) |> 
  janitor::clean_names() |> 
  mutate(
     sex = case_match(
       sex,
        1 ~ "male",
       2 ~ "female"
     ),
     education = case_match(
        education,
       1 ~ "Less than high school",
       2 ~ "High school equivalent",
       3 ~ "More than high school"
     )
  ) |> 
  filter( age >= "21", bmi != "NA", education != "NA", sex != "NA")

```



